\stdpopsim: a community toolkit for population genetic simulation

### 2019-12-07 popSimManu.tex
cleaning after past PR

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution from PopSim currently is \stdpopsim, a community
maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about its genome, with information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically
observed heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of known maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the HapMapII (\cite{international2007second}) and
Decode (\cite{kong2010fine}) genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios.

@@

\subsection*{Workflow for two population analysis}
Our goal for the two population analysis pipeline was to explore two population inference
on some of the multi-population demographic models implemented in \stdpopsim.
The models we examined include the human Gutenkunst model (\cite{gutenkunst2009inferring}) as
well as the \emph{Drosophila} Li & Stephan model (\cite{li2006inferring}). The human
model is a three population model (Africa, Europe, and Asia) including post-divergence
migration and exponential growth (Figure \ref{fig:IM_popn_human}C), whereas the
\emph{Drosophila} model is a two population model (Africa and Europe) with no post-divergence
migration and constant population sizes (Figure Sx).

@@

For both \dadi and \fastsimcoal2, our general approach was to fit a two population
isolation with migration (IM) model with constant population sizes to the simulated
data from the human and \emph{Drosophila} models. This IM model contains six parameters:
the ancestral population size, the size of population 1 after the split, the size of
population 2 after the split, the divergence time, and two migration rate parameters.
Importantly, this meant that for both the human and \emph{Drosophila} inference, the
fitted model did not match the simulated model (Figure \ref{fig:IM_popn_human}C and Sx).
In the human case, we therefore performed inference solely on the Africa
and Europe populations, meaning that the Asia population functioned as a ghost
population that was ignored by our inference. Our motivation for fitting this simple
IM model was to mimic the typical approach of two population inference on empirical
data, where the user is not aware of the 'true' underlying demography and the inference
model is often misspecified. To ground-truth our inference approach, we also conducted
inference on a generic IM model that was identical to the model used for inference.

To perform inference, we began by running simulations under each model, running
three replicates to inspect for inter-replicate variability. For humans, we ran
whole genome simulations (chromosomes 1-22) under the Gutenkunst model (\cite{gutenkunst2009inferring}),
using the HapMap II recombination map (\cite{international2007second}) and taking 20
samples each from the Europe and Africa populations. For \emph{Drosophila},
we simulated 3 Mb of chromosome 2R under the Li & Stephan model (\cite{li2006inferring}),
using the map inferred by \cite{comeron2012many} and taking 50 samples from each of
the Africa and Europe populations. We tried to run whole genome simulations under
this model but after XXXX amount of time... (JRA fill in details here). For the generic
IM simulations, we used the human genome (chromosomes 1-22) and HapMap II recombination
map (\cite{international2007second}) and took 20 samples per population.

Following simulation, we outputted tree sequences and masked low recombination
regions using the same approach described for the single population pipeline above. We
converted tree sequences into a two dimensional site frequency spectrum for all
chromosomes in the appropriate format for \dadi and \fastsimcoal2. For each simulation
replicate, we performed 10 runs of \dadi and \fastsimcoal2 and checked for convergence.
Detailed settings for \dadi and \fastsimcoal2 can be found in the Snakefile
on the popgensims/analysis GitHub repository (https://github.com/popgensims/analysis). The parameter
estimates from the top run from each simulation replicate are shown in Figures \ref{fig:IM_popn_human}C and Sx.

@@

To visualize our results, we plotted the inferred population size trajectories
for each simulation replicate along side the true (census) population sizes
(Figures \ref{fig:IM_popn_human}C Sx). Here, unlike the single population pipeline,
 we focus on census size rather than the inverse coalescence rate as the 'true' population size.

>>>>

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution from PopSim currently is \stdpopsim, a community
maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about its genome, with information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically
observed heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of known maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the HapMapII 
\cite{international2007second}) and
Decode (\cite{kong2010fine}) genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios.

@@

\subsection*{Workflow for two population analysis}
Our goal for the two population analysis pipeline was to explore two population inference
on some of the multi-population demographic models implemented in \stdpopsim.
The models we examined include the human Gutenkunst model (\cite{gutenkunst2009inferring}) as
well as the \emph{Drosophila} Li and Stephan model (\cite{li2006inferring}). The human
model is a three population model (Africa, Europe, and Asia) including post-divergence
migration and exponential growth (Figure \ref{fig:IM_popn_human}C), whereas the
\emph{Drosophila} model is a two population model (Africa and Europe) with no post-divergence
migration and constant population sizes (Figure \ref{fig:two_popn_fly}).

@@

For both \dadi and \fastsimcoal2, our general approach was to fit a two population
isolation with migration (IM) model with constant population sizes to the simulated
data from the human and \emph{Drosophila} models. This IM model contains six parameters:
the ancestral population size, the size of population 1 after the split, the size of
population 2 after the split, the divergence time, and two migration rate parameters.
Importantly, this meant that for both the human and \emph{Drosophila} inference, the
fitted model did not match the simulated model (Figure \ref{fig:IM_popn_human}C and Sx).
In the human case, we therefore performed inference solely on the Africa
and Europe populations, meaning that the Asia population functioned as a ghost
population that was ignored by our inference. Our motivation for fitting this simple
IM model was to mimic the typical approach of two population inference on empirical
data, where the user is not aware of the `true' underlying demography and the inference
model is often misspecified. To ground-truth our inference approach, we also conducted
inference on a generic IM model that was identical to the model used for inference.

To perform inference, we began by running simulations under each model, running
three replicates to inspect for inter-replicate variability. For humans, we ran
whole genome simulations (chromosomes 1-22) under the Gutenkunst model (\cite{gutenkunst2009inferring}),
using the HapMap II recombination map (\cite{international2007second}) and taking 20
samples each from the Europe and Africa populations. For \emph{Drosophila},
we simulated 3 Mb of chromosome 2R under the Li and Stephan model (\cite{li2006inferring}),
using the map inferred by \cite{comeron2012many} and taking 50 samples from each of
the Africa and Europe populations. We tried to run whole genome simulations under
this model but after XXXX amount of time... (JRA fill in details here). For the generic
IM simulations, we used the human genome (chromosomes 1-22) and HapMap II recombination
map (\cite{international2007second}) and took 20 samples per population.

Following simulation, we outputted tree sequences and masked low recombination
regions using the same approach described for the single population pipeline above. We
converted tree sequences into a two dimensional site frequency spectrum for all
chromosomes in the appropriate format for \dadi and \fastsimcoal2. For each simulation
replicate, we performed 10 runs of \dadi and \fastsimcoal2 and checked for convergence.
Detailed settings for \dadi and \fastsimcoal2 can be found in the Snakefile
on the popgensims/analysis GitHub repository (\url{https://github.com/popgensims/analysis}).
The parameter estimates from the top run from each simulation replicate
are shown in Figures \ref{fig:IM_popn_human}C and Sx.

@@

To visualize our results, we plotted the inferred population size trajectories
for each simulation replicate along side the true (census) population sizes
(Figures \ref{fig:IM_popn_human}C Sx). Here, unlike the single population pipeline,
we focus on census size rather than the inverse coalescence rate as the `true' population size.
